Neural Power Units

### 2020-12-17 paper/main.tex
rename table...
Merge arxiv branch
Includes slides / fixed references

\begin{abstract}
  Conventional Neural Networks can approximate simple arithmetic operations,
  but fail to generalize beyond the range of numbers that were seen during
  training.  \emph{Neural Arithmetic Units} aim to overcome this difficulty,
  but current arithmetic units are either limited to operate on positive
  numbers or can only represent a subset of arithmetic operations. We introduce
  the \emph{Neural Power Unit} (NPU)\footnote{
    Implementation of Neural Arithmetic Units: \url{github.com/nmheim/NeuralArithmetic.jl}.
    The code to reproduce our experiments is available at \url{github.com/nmheim/NeuralPowerUnits}.
  }
  that operates on the \emph{full domain of
  real numbers} $\mathbb{R}$ and is capable of \emph{learning arbitrary power functions} in
  a single layer.  The NPU thus fixes the shortcomings of existing arithmetic
  units and extends their expressivity. We achieve this by using
  complex arithmetic without requiring a conversion of the network to complex
  numbers $\mathbb{C}$.
  A simplification of the unit to the \emph{RealNPU} yields a highly transparent model.
  We show that the NPUs outperform their competitors in terms of
  accuracy and sparsity on artificial arithmetic datasets, and that the RealNPU can discover
  the governing equations of a dynamical system only from data.
\end{abstract}

@@

If an ODE model is composed of dense layers, its direct interpretation is
problematic and has to be performed retrospectively. The class of models based on
SINDy is transparent by design, however it can only provide explanation
within a linear combination of predefined set of basis functions. Thus, it
cannot learn models with unknown fractional powers. 
With this experiment we aim to show that the NPU can \emph{potentially} discover
exact ODE models\footnote{The demonstration given here is not intended to be
used in practice. For real-world predictions in such a sensitive area much more
post-processing is needed to ensure safe predictions.}.
An example of an ODE that contains powers is a modification of the well-known
epidemiological SIR model \citep{kermack_contribution_1927} to fractional powers
(fSIR, \cite{taghvaei_fractional_2020}), which was shown to be a beneficial
modification for modelling the COVID-19 outbreak.  The SIR model is
built from three variables: $S$ (susceptible), $I$ (infectious), and $R$
(recovered/removed).  Arguably the most important part of the model is the
transmission rate $r$, which is typically taken to be proportional to the
product of $S$ and $I$.  \citet{taghvaei_fractional_2020} argue that,
especially in the initial phase of an epidemic, the boundary areas of infected
and susceptible cells scale with a fractional power, which leads to
Eq.~\ref{eq:fracsir_r}:
\begin{align}
  \label{eq:fracsir}
  \frac{dS}{dt} &= -r(t) + \eta R(t), &
  \frac{dI}{dt} &=  r(t) - \alpha I(t), &
  \frac{dR}{dt} &= \alpha I(t) - \eta R(t),\\
  \label{eq:fracsir_r}
  & & r(t) &= \beta I(t)^\gamma S(t)^\kappa,
\end{align}

@@

\section{Statement of Broader Impact}%
\label{sec:statement_of_broader_impact}

@@

\section{Acknowledgements}%
\label{sec:acknowledgements}

>>>>

\begin{abstract}
  Conventional Neural Networks can approximate simple arithmetic operations,
  but fail to generalize beyond the range of numbers that were seen during
  training.  \emph{Neural Arithmetic Units} aim to overcome this difficulty,
  but current arithmetic units are either limited to operate on positive
  numbers or can only represent a subset of arithmetic operations. We introduce
  the \emph{Neural Power Unit} (NPU).\footnote{
    Implementation of Neural Arithmetic Units: \url{github.com/nmheim/NeuralArithmetic.jl}
    The code to reproduce our experiments is available at \url{github.com/nmheim/NeuralPowerUnits}.
  }
  that operates on the \emph{full domain of
  real numbers} $\mathbb{R}$ and is capable of \emph{learning arbitrary power functions} in
  a single layer.  The NPU thus fixes the shortcomings of existing arithmetic
  units and extends their expressivity. We achieve this by using
  complex arithmetic without requiring a conversion of the network to complex
  numbers $\mathbb{C}$.
  A simplification of the unit to the \emph{RealNPU} yields a highly transparent model.
  We show that the NPUs outperform their competitors in terms of
  accuracy and sparsity on artificial arithmetic datasets, and that the RealNPU can discover
  the governing equations of a dynamical system only from data.
\end{abstract}

@@

If an ODE model is composed of dense layers, its direct interpretation is
problematic and has to be performed retrospectively. The class of models based on
SINDy is transparent by design, however it can only provide explanation
within a linear combination of predefined set of basis functions. Thus, it
cannot learn models with unknown fractional powers. 
With this experiment we aim to show that the NPU can \emph{potentially} discover
exact ODE models.\footnote{The demonstration given here is not intended to be
used in practice. For real-world predictions in such a sensitive area much more
post-processing is needed to ensure safe predictions.}
An example of an ODE that contains powers is a modification of the well-known
epidemiological SIR model \citep{kermack_contribution_1927} to fractional powers
(fSIR, \cite{taghvaei_fractional_2020}), which was shown to be a beneficial
modification for modelling the COVID-19 outbreak.  The SIR model is
built from three variables: $S$ (susceptible), $I$ (infectious), and $R$
(recovered/removed).  Arguably the most important part of the model is the
transmission rate $r$, which is typically taken to be proportional to the
product of $S$ and $I$.  \citet{taghvaei_fractional_2020} argue that,
especially in the initial phase of an epidemic, the boundary areas of infected
and susceptible cells scale with a fractional power, which leads to
Eq.~\ref{eq:fracsir_r}:
\begin{align}
  \label{eq:fracsir}
  \frac{dS}{dt} &= -r(t) + \eta R(t), &
  \frac{dI}{dt} &=  r(t) - \alpha I(t), &
  \frac{dR}{dt} &= \alpha I(t) - \eta R(t),\\
  \label{eq:fracsir_r}
  & & r(t) &= \beta I(t)^\gamma S(t)^\kappa,
\end{align}

@@

\section*{Broader Impact}%
\label{sec:statement_of_broader_impact}

@@

\section*{Acknowledgements and Disclosure of Funding}%
\label{sec:acknowledgements}
